Symmetric Peer-to-Peer Applications

### 2022-08-19 tpd-22.tex
...

Collaborative networked applications allow multiple users to interact
remotely and yet share the same experience in real time.
Examples of such \emph{symmetric distributed applications}~\cite{gals}
are shared documents, watch parties, and multi-player games.
%
In order to reproduce the exact behavior in multiple devices, the application
must be able to synchronize its time and execution across the network.
In addition, since users can interact asynchronously with the application,
event occurrences must somehow be synchronized back across devices.

@@

Croquet%
\footnote{Croquet.io: https://croquet.io/}~\cite{TODO} guarantees
bit-identical real-time behavior for every user in a collaborative
distributed environment.
%
As detailed in Figure~\ref{fig.algos}, a centralized \emph{reflector} issues
periodic ticks, such that all nodes in the network advance together according
to a synchronized global shared clock.
If an user generates an event, it is delayed and sent back to the reflector,
which broadcasts the event in the next tick, which all nodes apply in sync
(including the originating node).
%
Croquet takes periodic snapshots of the whole application state in order to
support late joins to a running session.
The new node just needs to restore the latest snapshot and simulate the
remaining events to reach the current running state.
%
As indicated in Figure~\ref{fig.table}, Croquet relies on a centralized
network, in which nodes advance in sync, and in which event outcomes depend
on the originating node to be online.

GALS~\cite{TODO} shares the same goals with Croquet, but with some tradeoffs,
mostly favoring clients with slow connections.
As detailed in Figure~\ref{fig.algos}, instead of advancing ticks in sync with
the server, clients have their own local clocks.
Also, event generation is delayed dynamically according to the slowest client.
%
On the one hand, ticks do not generate any traffic and clients have smooth
frame transitions, even those with poor connections.
On the other hand, event generation requires two round trips to the server and
clients may experience occasional freezes.
The syncing protocol also requires bookkeeping to deal with clock drifts and
client disconnections.
%
As indicated in Figure~\ref{fig.table}, GALS relies on a centralized
network, in which nodes advance independently, and in which event outcomes
depend on the originating node to be online.

@@

An antagonistic approach to deterministic reactions to events is to model the
application data structures with conflict-free replicated data types
(CRDTs)~\cite{TODO}.
CRDTs are designed in such a way that all operations are commutative, so that
the order in which they are applied does not affect the final outcome.
%
As detailed in Figure~\ref{fig.algos}, peers can communicate operations
directly to each other with no central authority.
Also, since operations need not to be ordered, they can be applied at the
very first moment they are generated or received, even if the peers are
offline.
%
Since operations must be commutative, it is not possible to timestamp them
in a global shared clock.
Hence, there is no notion of a timeline with peers going through bit-identical
steps, but they do eventually reach the same state.
%
On the one hand, the main advantage of CRDTs is they support the concept of
local-first software~\cite{TODO}, since they work offline.
On the other hand, they provide very restrictive APIs (the CRDTs themselves),
and do not support real-time consensus among peers.
%
Automerge~\cite{TODO} along with its accompanying Hypermerge peer-to-peer
infrastructure is an example of an implementation of this approach.

In this work, we try to .

- offline, local first
- DAG, no consensus, partial order
    - branches in parallel w/ T,T+d not required to be applied in order

- all still deterministic, stateless

Croquet OS is a.
- centralized
    - synchronous
        - croquet
	- PALS
    - asynchronous
        - gals
- decentralized
    - CRDTs (asynchronous) (not realtime) (pause would not be bit identical)
    - rollback (async)

- local-first + CRDTs + pushpin

- time machine
    - paper que eu achei ha um tempo
    - braid

@@

\bibliographystyle{IEEEtran}
%\bibliography{tpd-22}

>>>>

\IEEEPARstart{C}{ollaborative} networked applications allow multiple users to
interact remotely and yet share the same experience in real time.
Examples of such \emph{symmetric distributed applications}~\cite{gals}
are shared documents, watch parties, and multi-player games.
%
In order to reproduce the exact behavior in multiple devices, the application
must be able to synchronize its time and execution across the network.
In addition, since users can interact asynchronously with the application,
event occurrences must somehow be synchronized back across devices.

@@

Croquet%
\footnote{Croquet.io: https://croquet.io/}~\cite{croquet} guarantees
bit-identical real-time behavior for every user in a collaborative
distributed environment.
%
As detailed in Figure~\ref{fig.algos}, a centralized \emph{reflector} issues
periodic ticks, such that all nodes in the network advance together according
to a synchronized global shared clock.
If an user generates an event, it is delayed and sent back to the reflector,
which broadcasts the event in the next tick, which all nodes apply in sync
(including the originating node).
%
Croquet takes periodic snapshots of the whole application state in order to
support late joins to a running session.
The new node just needs to restore the latest snapshot and simulate the
remaining events to reach the current running state.
%
As indicated in Figure~\ref{fig.table}, Croquet relies on a centralized
network, in which nodes advance in sync, and in which event outcomes depend
on the originating node to be online.

GALS~\cite{gals} shares the same goals with Croquet, but with some tradeoffs,
mostly favoring clients with slow connections.
As detailed in Figure~\ref{fig.algos}, instead of advancing ticks in sync with
the server, clients have their own local clocks.
Also, event generation is delayed dynamically according to the slowest client.
%
On the one hand, ticks do not generate any traffic and clients have smooth
frame transitions, even those with poor connections.
On the other hand, event generation requires two round trips to the server and
clients may experience occasional freezes.
The syncing protocol also requires bookkeeping to deal with clock drifts and
client disconnections.
%
As indicated in Figure~\ref{fig.table}, GALS relies on a centralized
network, in which nodes advance independently, and in which event outcomes
depend on the originating node to be online.

@@

An antagonistic approach to deterministic reactions to events is to model the
application data structures with conflict-free replicated data types
(CRDTs)~\cite{crdts}.
CRDTs are designed in such a way that all operations are commutative, so that
the order in which they are applied does not affect the final outcome.
%
As detailed in Figure~\ref{fig.algos}, peers can communicate operations
directly to each other with no central authority.
Also, since operations need not to be ordered, they can be applied at the
very first moment they are generated or received, even if the peers are
offline.
%
Since operations must be commutative, it is not possible to timestamp them
in a global shared clock.
Hence, there is no notion of a timeline with peers going through bit-identical
steps, but they do eventually reach the same state.
%
On the one hand, the main advantage of CRDTs is they support the concept of
local-first software~\cite{local}, since they work offline.
On the other hand, they provide very restrictive APIs (the CRDTs themselves),
and do not support real-time consensus among peers.
%
As indicated in Figure~\ref{fig.table}, CRDTs work on decentralized networks,
in which nodes advance independently, and in which operations can be applied
immediately, even while offline.
%
Automerge along with its accompanying Hypermerge peer-to-peer infrastructure
is an example of an implementation of this approach~\cite{p2p.automerge,p2p.pushpin}.

In this work, as indicated in Figure~\ref{fig.table}, our goal is to support
peer-to-peer real-time symmetric execution, while still tolerating offline and
out-of-order event generation.
%
As detailed in Figure~\ref{fig.algos}, peers have independent timelines and
can communicate events directly to each other.
Events are delayed to be able to reach other peers in time.
In the case a deadline is missed, the peer rolls back in time and fast
forwards execution until it reaches real time again.
%
Like Croquet and GALS, peers compute events in sequence using only
deterministic and stateless calls.

TODO
- photon unreal
- time machine
    - paper que eu achei ha um tempo
    - braid

@@

\bibliographystyle{IEEEtran}
\bibliography{tpd-22}
